\frontmatter
\title{\gl}
\author{Sven C. Vogel
\thanks{This work could not have been possible without the efforts of Drs Allen C. Larson and Robert B. Von Dreele to create and maintain GSAS. I am indebted to Nina J. Lane (Drexel University) for the majority of the effort to generate this manual - you wouldn't read this without her! The Lujan Neutron Scattering Center at the Los Alamos Neutron Science Center is funded by the Department of Energy's Office of Basic Energy Science.  The Los Alamos National Laboratory is operated by the Los Alamos National Security LLC under the DOE Contract of DE-AC52-06NA25396.}\\
Manuel Lujan Jr. Neutron Scattering Center\\ Los Alamos National Laboratory, Los Alamos, New Mexico, 87545, U.S.A.\\

{sven@lanl.gov}
}
\date{December 2010\thanks{First drafted September 2009}\\
\vspace{0.25in}
If you used gsaslanguage and found it helpful, please reference this paper:\\
\vspace{0.25in}
S. C. Vogel,\\
\emph{gsaslanguage: a GSAS script language for automated Rietveld refinements of diffraction data},\\
J. Appl. Cryst. (2011). 44, 873-877\\
\url{http://dx.doi.org/10.1107/S0021889811023181}}
\begin{titlingpage} 
  \maketitle
\end{titlingpage}
\mainmatter
\begin{abstract} 
The commands of the \gl\hspace{0.5em}software are documented. The software utilizes the GSAS package and is intended to simplify automation of the analysis of large numbers of datasets, document refinement strategies and foster efficient exchange of refinement strategies between users and beam line scientists. Use of the bash shell and standard Unix text processing tools, available natively on Linux and Mac OSX platforms and via the free cygwin software on Windows systems, make this software platform independent.
\end{abstract}
\newpage
\tableofcontents
